Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-04-02 Ben_Manuscripts/thesis/intro.tex
Calculate local density

\chapter{Introduction}\label{chapter:intro}

@@

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers.~\cite{goosen_fouling_2005}
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.~\cite{werber_materials_2016}
    \item The diversity in size and chemical topology of contaminants makes the
    design of membrane processes meant to separate these components complex.
    \item To deal with complex aqueous streams, and to prevent excessive membrane
    fouling, multiple types of membranes are often used in series, removing 
    larger particles first.~\cite{suk_membranebased_2006}
  \end{itemize}

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Modern membrane design is completely dependent on the target particle 
  separation. 
  \begin{itemize}
    \item Separation membranes are most broadly classified based on the size of 
    the particle they reject.
    \item Microfiltration (MF) membranes contain pores with diameters ranging from 
    100-10000 nm. They can separate large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration (UF) membranes have pore diameters of about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
    \item Nanofiltration (NF) membranes have pores on the order of 1 nm in diameter and can 
    be used for the separation of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    \item Reverse osmosis (RO) membranes are dense amorphous polymers with tortuous 
    pathways formed by collapsed polymer networks. Their dense polymer architecture 
    rejects all solutes to a high degree except for water making them useful for 
    separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of the different classes of membrane technologies in 
    Figure~\ref{fig:size_regimes}.
  \end{itemize}

@@

  Membranes are generally designed to optimize two key performance metrics: 
  permeability and selectivity.~\cite{geise_water_2011}
  %BJC2: Not sure the right place to put citation since its all described in the above citation. I could add more, but seems pointless.
  \begin{itemize}
    \item Permeability is used to quantify the rate at which solvent and solutes
    pass through the membrane.
    \item Selectivity quantifies the relative rates of passage of two substances.
    \item A related quantity to selectivity, rejection, quantifies the fraction
    of undesired solute that does not pass through the membrane.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.~\cite{werber_materials_2016}
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi \textit{et al.} calculated that a theoretical tripling of membrane permeability
    would only reduce energy consumption by up to 15\% in a seawater RO plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.~\cite{singh_production_2009}
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment.~\cite{shannon_science_2009,fritzmann_state---art_2007}
%    \item The well-known permeability-selectivity tradeoff, common to dense polymer 
%    membranes, directly correlates an increase in water permeability with an 
%    increase in solute permeability.~\cite{geise_water_2011}
    \item Addressing these problems with highly selective separations may lead to 
    more gains.~\cite{deshmukh_desalination_2015}
 \end{itemize}

@@

  Selective separations are useful in a wide range of applications. They can be used
  to separate components in both gaseous and aqueous phases.~\cite{baker_gas_2014}. We 
  will focus this section on aqueous separations since they are the target process
  for the following studies.

  \textit{Desalination}: Creating potable water from seawater or brackish water is of 
  paramount importance in water-scarce areas. 
  \begin{itemize}
    \item It is estimated that by 2025, 1.8 billion people may live in water-stressed
    regions.~\cite{navarro-ortega_managing_2015} %There's a water stress index in fritzmann_2007
    \item Membrane deslination via RO and thermal muli-stage flash distillation are 
    both widely used processes for creating potable water in these regions.~\cite{fritzmann_state---art_2007}
  	\item Compared with thermal distillation techniques, RO has been shown to 
  	be more environmentally friendly and economical due to its lower energy
  	requirements.\cite{morton_environmental_1997}
% 	\item In general, RO has a smaller footprint % BJC: need citation
  	\item However, thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.\cite{bhojwani_technology_2019}
  \end{itemize}

  \textit{Organic Micropollutants}: Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which can have adverse effects on human health even at low 
  concentrations.\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceuticals, personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply.~\cite{barbosa_occurrence_2016}
    \item Sources of micropollutants incude industrial wastewater, 
    agricultural runoff, landfill leachates and domestic effluents.~\cite{mompelat_occurrence_2009}
    %BJC2: There is a review I need to wade through to bulk this up.
    \item Unfortunately, most waste water treatment plants are not 
    designed for complete removal of these contaminants.~\cite{tijani_review_2013}
    \item There is a need for a new generation of membranes that can better
    target specific molecules.
  \end{itemize}

@@

  Municipal waste-waters are rich in carbon, nitrogen and phosphorus-containing 
  compounds.~\cite{romero_raw_2013}
  \begin{itemize}  
    \item The recovery of such products, which can be achieved using a 
    selective membrane, has numerous potential uses.\cite{sales_resource_2015}
    \item For example, nitrogen and phosphorus recovery can help sustain fertilizer 
    production which will help meet global food demand as population continues
    to increase.\cite{xie_membrane-based_2016}
  \end{itemize}

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater.\cite{werber_materials_2016} 
  \begin{itemize}
    \item For example, flowback water, produced during hydraulic fracturing of
    shale formations consists of relatively high concentrations of salts, metals,
    and soluble organic compounds.~\cite{orem_organic_2014}
    \item The majority of this water is disposed through deep well injection, 
    however there is a growing public concern about its management which has 
    prompted the use of separation technologies such as RO and NF in order to 
    reduce the volume of contaminated water.~\cite{gregory_water_2011}
    \item Some of the dissolved organic compounds in flowback water, such as acetate, 
    are potentially valuable and can be recovered with highly selective membranes
    \cite{dischinger_application_2017}.
  \end{itemize}

%BJC: this is more of a gas separation
%  \textit{Breathable barriers}:
%  
%  Finally, there is a great deal of military interest in creating breathable 
%  barriers which selectively allow the passage of water in both directions but
%  blocks out harmful contaminants.
%  \begin{itemize}
%    \item Need to allow sweating
%    \item Can even put catalytic groups in the pores to break down contaminants
%  \end{itemize}

  \section{Historical Development of Membrane Technology}
  %BJC2: maybe better to title this section "Historical Development of Membrane Technology"

  Membranes for separations are a relatively young technology. Over the past 60 years, 
  the field has grown into a commercially viable technology and the focus has shifted
  primarily towards incremental improvements in performance and most importantly,
  targeted separations. This section will briefly cover a subset of the most relevant
  membrane technologies which have ultimately inspired this thesis.

@@

  The first aqueous separations membranes were developed for RO applications 
  in the 1960s and were made of the amorphous polymer cellulose acetate (CA).~\cite{kesting_semipermeable_1965}
  \begin{itemize}
    \item Reid and Spencer were among the first to show the ability of
    these membranes to reject NaCl to high degrees ($\sim$ 99 \%) while subject 
    to high applied pressures (270 atm).~\cite{reid_ultrafiltration_1960}
    \item The most limiting drawback to CA membranes is their chemical instability, 
    undergoing severe hydrolysis at pH values outside the range 4--6.~\cite{vos_kinetic_1966}
  \end{itemize}

  Amorphous thin film composite (TFC) polyamide membranes are the current industry 
  standard for high purity RO separations.~\cite{baker_membrane_2012}
  \begin{itemize}  
    \item Developed in the 1980s, polyamide membranes are capable of rejecting salt
    to a higher degree and can operate at wider pH ranges and higher feed pressures than 
    CA membranes.~\cite{cadotte_interfacial_1981}
    \item TFC polyamide membranes are typically formed by the interfacial polymerization
    of a polyamide selective layer on the surface of a microporous polysulfone support 
    layer.    
    \item The support layer gives the membrane mechanical strength without hindering the
    flow of water thereby permitting a very thin selective layer where molecular 
    separations occur.~\cite{kucera_biofouling_2019}
  \end{itemize}

  The dense polyamide selective layers of TFC membranes require high feed pressures,
  in the range of 5 -- 120 bar, in order to achieve a useful permeate flux.~\cite{van_der_bruggen_review_2003}
  \begin{itemize}
    \item While the energy requirements to attain these pressures are still less than 
    the energy requirements of thermal distillation\cite{morton_environmental_1997}, it 
    is often desirable to create membranes that can operate at lower feed pressures.
    %BJC2: probably could use some citations here
    \item It is not always necessary to achieve the high purity permeate streams
    resulting from RO.
    \item It is also possible that a target separation may be possible with 
    membranes containing larger pores.
    \item This has driven research into nanofiltration.
  \end{itemize}

@@

  Since their introduction in the late 1980s, NF membranes have been applied towards
  an increasing number of separations since they bridge the gap in capabilities
  between reverse osmosis and ultrafiltration.~\cite{eriksson_nanofiltration_1988}
  \begin{itemize}
    \item Nanofiltration membranes have pores on the order of 1 nm in size. 
    \item Larger pore sizes allow lower operating pressures (3 -- 20 bar) which lowers 
    energy consumption.~\cite{van_der_bruggen_review_2003}
    \item An ideal NF membrane should have densely packed, uniform-sized and non-tortuous pores.~\cite{jackson_nanoporous_2010}
    This combination has been practically difficult to realize.
  \end{itemize}

  The most widely used methods of NF membrane fabrication are interfacial polymerization
  and phase inversion.
  \begin{itemize}
    \item NF membranes created by interfacial polymerization are made similarly to
    RO membranes. By modifying the monomer materials, one can control the structural morphology
    in order to create pores larger than those in RO membranes.~\cite{seman_nanofiltration_2010}
    \item During phase inversion, one submerges a polymer, dissolved in a solvent, in
    a non-solvent. A solid, porous polymer membrane is all that remains once all
    solvent has been removed by non-solvent exchange.~\cite{smolders_microstructures_1992}
    \item Although commercially dominant, both of these techniques suffer from
    tortuous pores and polydisperse pore sizes which results in far from ideal 
    selectivities.
  \end{itemize}

  Track-etching is another NF fabrication technique which attempts to eliminate 
  polydispersity and tortuosity. During track-etching, a polymer film is bombarded with 
  charged particles, then chemically etched to create pores.~\cite{apel_track_2001}
  \begin{itemize}
	\item Although the pores are uniform and non-tortuous, benefiting selectivity, the 
	membranes have an extremely low pore density and subsequently low permeability. 
  \end{itemize}

@@

  %BJC2: would a figure showing the nanostructures go well here? (a graphene film, a carbon nanotube, a MOF, a zeolite)

  %BJC2: Each tecnhology below will likely have its own heading and be split into 2 or more 
  % paragraphs.
  \textit{Graphene and graphene oxide} Ultrathin-film graphene and graphene oxide membranes
  are an active area of research because they are atomically thick materials and therefore 
  offer potential for extremely high permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}
    \item The tightly packed aromatic network of $sp^2$ hybridized carbon atoms that 
    make up graphene sheets precludes the passage of even the smallest monoatomic species, helium.~\cite{bunch_impermeable_2008}
    \item Controllably-sized pores have been created in pristine graphene by plasma etching
    and by charged particle bombardment.~\cite{surwade_water_2015,russo_atom-by-atom_2012}
    \item Graphene oxide (GO), a highly oxidized modification of graphene, consist of epoxy,
    hydroxyl and carboxylate functional groups.~\cite{hegab_graphene_2015}
    \item Many groups study GO membranes because they have improved antibacterial properties 
    over graphene, leading to less membrane fouling, and because their functionality can be
    modified in order to achieve solute-specific separations.~\cite{liu_antibacterial_2011,he_bioinspired_2013}
    \item Despite their promise, scalable synthesis of monolayer graphene and GO membranes without
    introducing microscopic, performance degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
    \item Multilayered GO membranes have shown considerably more progress since their 
    synthesis procedures are low cost and scalable.~\cite{zhang_size-controlled_2009}
    \item GO sheets are particularly easy to make into multilayered membranes since they 
    stack on top of each other, held together by hydrogen bonds.~\cite{homaeigohar_graphene_2017}
    \item The potential of graphene and GO membrane certainly merits their continued exploration. Many
    groups are still working in order to work on better control and standardization of their synthesis,
    improving their designability for specific separations, and further improving their scalability.~\cite{wei_multilayered_2018}
  \end{itemize}

  \textit{Carbon nanotubes}: Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011}
  \begin{itemize}
    \item CNTs are an allotrope of carbon consisting of tubular graphitic layers.~\cite{iijima_helical_1991}
    \item One can control CNT diameters down to 0.5 nm in size which has inspired the
    creation of CNT membranes for reverse osmosis applications.~\cite{sun_creating_2000}
    \item Molecular dynamics simulations have predicted and experiments have confirmed that
    CNTs are extremely fast transporters of water.~\cite{hummer_water_2001,joseph_why_2008}
    \item MD simulations by Kalra \textit{et al.} predicted water transport at rates comparable
    to the transmembrane protein aquaporin-1.~\cite{kalra_osmotic_2003}
    \item Experiments have exhibited water transport rates orders of magnitude higher than 
    predictions by continuum modeling.~\cite{holt_fast_2006}
    \item This enhancement has been explained as a result of nearly friction-less single-file water
    flow through the hydrophobic CNT interiors.~\cite{kalra_osmotic_2003}
    \item Functionalization of the CNT pore entrances can potentially lead to highly selective separations.~\cite{fornasiero_ion_2008}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them,
    however facile incorporation at scale remains a challenge yet to be overcome.~\cite{kim_polysulfone_2007}
  \end{itemize}

  \textit{Zeolites}: Zeolite membranes offer the potential for permeabilities comparable to ultrafiltration, 
  with selectivities as good as NF and RO.~\cite{hoek_nanotechnology-based_2014}
  \begin{itemize}  
    \item Zeolites have highly uniform nm-sized crystalline structures with cage-like
    cavities that allow movement and trapping of small solutes. 
    \item The crystalline frameworks are typically formed by networks of silicon 
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
    \item One can replace the silicon and aluminum atoms via ion exchange in order
    to control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}
    \item Zeolite-coated ceramic membranes have shown the most promise for RO applications
    because of their mechanical stability under high applied pressure and chemical
    stability for withstanding disinfectants.~\cite{pendergast_review_2011}
    \item A number of studies have tested the permeability and sodium salt rejection
    of various zeolite membranes, however their high permeabilities typcially come
    at the cost of low selectivities and vice versa.~\cite{li_desalination_2004,duke_seawater_2009}
    \item Zeolites are also prone to defects in the crystalline structure which 
    would require substantial modification to the traditional synthesis procedures
    in order to fix.~\cite{kumakiri_application_2000}
  \end{itemize}

  \textit{Metal organic frameworks}: Finally, metal organic frameworks (MOFs) have been
  recently introduced to the field of aqueous separations.
  \begin{itemize}
    \item MOFs are highly porous and flexible crystalline structures made of metal
    clusters coordinated to organic ligands.~\cite{furukawa_chemistry_2013}
    \item There is a much wider breadth of research into MOFs as gas separation membranes
    because they tend to be unstable in water.~\cite{feng_construction_2013}
    \item But a new generation of water-stable MOFs is making it possible to use them
    in aqueous separations.~\cite{wang_ultrastable_2015}
    \item Purification of both aqueous and gaseous streams using MOFs are 
    typically achieved by selective adsorption or catalytic degradation.~\cite{li_metalorganic_2012}    
    \item MOFs have been applied towards the removal of a number of common water 
    contaminants including dyes, pesticides, endocrine disruptors, and pharmaceuticals.
    \cite{zhang_fabrication_2015,seo_adsorptive_2015,li_metalorganic_2015,hasan_adsorptive_2012}
    \item Although aqueous MOF separations are theorized to surpass the performance 
    of conventional membrane separations, it is still a budding field. The stability and
    reusability of the material must be guaranteed before they can be implemented
    at scale.~\cite{dias_towards_2015}
  \end{itemize}

@@

  Amphiphilic molecules that self-assemble to form ordered nanostructures have shown
  potential as nanoporous membranes.
  \begin{itemize}
    \item People are particularly interested in self-assembled materials because they
    offer the potential of uniform-sized pores and tunable chemical and physical 
    properties.~\cite{werber_materials_2016}
    \item There are two major classes of self-assembled amphiphilic materials: block 
    copolymers and liquid crystals.
  \end{itemize}

>>>>

8\chapter{Introduction}\label{chapter:intro}

@@

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations. Untreated water can be a very complex solution
  with particles ranging in size from microns down to nanometers.~\cite{goosen_fouling_2005}
  Sediment, bacteria, algae, proteins, small organic molecules, and ions
  are all common components of aqueous streams.~\cite{werber_materials_2016}
  The diversity in size and chemical topology of contaminants makes the design of
  membrane processes meant to separate these components complex. To deal with 
  complex aqueous streams, and to prevent excessive membrane fouling, multiple 
  types of membranes are often used in series, removing larger particles 
  first.~\cite{suk_membranebased_2006}

  Modern membrane design is completely dependent on the target particle 
  separation. Separation membranes are most broadly classified based on the size of 
  the particle they reject. Microfiltration (MF) membranes contain pores with 
  diameters ranging from 100-10000 nm. They can separate large particles like 
  bacteria and protozoa.~\cite{ma_functionalized_2014} Ultrafiltration (UF) 
  membranes have pore diameters of about 5-500 nm and are useful for the separation
  of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
  Nanofiltration (NF) membranes have pores on the order of 1 nm in diameter and can 
  be used for the separation of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
  Reverse osmosis (RO) membranes are dense amorphous polymers with tortuous 
  pathways formed by collapsed polymer networks. Their dense polymer architecture 
  rejects all solutes to a high degree except for water making them useful for 
  separating hydrated ions from water.~\cite{warsinger_review_2018} We have 
  summarized the uses of the different classes of membrane technologies in 
  Figure~\ref{fig:size_regimes}.

@@

  Membranes are generally designed to optimize two key performance metrics: 
  permeability and selectivity.~\cite{geise_water_2011} Permeability is used to
  quantify the rate at which solvent and solutes pass through the membrane.
  Selectivity quantifies the relative rates of passage of two substances.
  A related quantity to selectivity, rejection, quantifies the fraction
  of undesired solute that does not pass through the membrane.

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.~\cite{werber_materials_2016} Increasing permeabilities
  beyond those already achieved may have only a small effect on energy 
  requirements and capital costs. Cohen-Tanugi \textit{et al.} calculated that 
  a theoretical tripling of membrane permeability would only reduce energy 
  consumption by up to 15\% in a seawater RO plant.~\cite{cohen-tanugi_quantifying_2014}
  Perhaps the most aggravating influence on energy consumption is the practical
  need for multiple membrane passes in order to achieve a desired purity.~\cite{singh_production_2009}
  Additionally, many membranes are not suited for high purity separations
  of small neutral species which necessitates chemical and capital costs for
  post-treatment.~\cite{shannon_science_2009,fritzmann_state---art_2007}
  Addressing these problems with highly selective separations may lead to 
  more signficant advances.~\cite{deshmukh_desalination_2015}

@@

  Selective separations are useful in a wide range of applications. They can be used
  to separate components in both gaseous and aqueous phases.~\cite{baker_gas_2014}. We 
  will focus this section on aqueous separations since they are the target process
  for the studies that follow.

  \textit{Desalination}: Creating potable water from seawater or brackish water is of 
  paramount importance in water-scarce areas. It is estimated that by 2025, 1.8 
  billion people may live in water-stressed regions.~\cite{navarro-ortega_managing_2015} %There's a water stress index in fritzmann_2007
  Membrane deslination via RO and thermal muli-stage flash distillation are 
  both widely used processes for creating potable water in these regions.~\cite{fritzmann_state---art_2007}
  Compared with thermal distillation techniques, RO has been shown to 
  be more environmentally friendly and economical due to its lower energy
  requirements.\cite{morton_environmental_1997} However, thermal techniques are still
  preferred where excess waste heat or cheap thermal energy is available, such as
  cogeneration plants.\cite{bhojwani_technology_2019}

  \textit{Organic Micropollutants}: Municipal and industrial wastewaters are contaminated
  by harmful micropollutants, which can have adverse effects on human health even at low 
  concentrations.\cite{schwarzenbach_challenge_2006} Micropollutants include pharmaceuticals,
  personal care products, hormones, pesticides and industrial chemicals which find their way
  into our drinking water supply.~\cite{barbosa_occurrence_2016} Sources of micropollutants
  incude industrial wastewater, agricultural runoff, landfill leachates and domestic 
  effluents.~\cite{mompelat_occurrence_2009} Unfortunately, most waste water treatment 
  plants are not designed for complete removal of these contaminants.~\cite{tijani_review_2013}
  There is a need for a new generation of membranes that can better target specific molecules.

@@

  Municipal waste-waters are rich in carbon, nitrogen and phosphorus-containing 
  compounds.~\cite{romero_raw_2013} The recovery of such products, which can be 
  achieved using selective membranes, has numerous potential uses.\cite{sales_resource_2015}
  For example, nitrogen and phosphorus recovery can help sustain fertilizer 
  production which will help meet global food demand as population continues
  to increase.\cite{xie_membrane-based_2016}

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater.\cite{werber_materials_2016} For example,
  flowback water, produced during hydraulic fracturing of shale formations consists
  of relatively high concentrations of salts, metals, and soluble organic 
  compounds.~\cite{orem_organic_2014} The majority of this water is disposed 
  through deep well injection, however there is a growing public concern about
  its management which has prompted the use of separation technologies such as 
  RO and NF in order to reduce the volume of contaminated water.~\cite{gregory_water_2011}
  Some of the dissolved organic compounds in flowback water, such as acetate, 
  are potentially valuable and can be recovered with highly selective membranes
  \cite{dischinger_application_2017}.

  \section{Historical Development of Membrane Technology}

  Membranes for separations are a relatively young technology that has expanded
  rapidly. Over the past 60 years, the field has grown into a commercially viable
  technology and the focus has shifted primarily towards incremental improvements
  in performance and most importantly, targeted separations. This section will 
  briefly cover a subset of the most relevant membrane technologies which have 
  ultimately inspired this thesis.

@@

  The first aqueous separations membranes were developed for RO applications 
  in the 1960s and were made of the amorphous polymer cellulose acetate (CA).~\cite{kesting_semipermeable_1965}
  Reid and Spencer were among the first to show the ability of these membranes
  to reject NaCl to high degrees ($\sim$ 99 \%) while subject to high applied 
  pressures (270 atm).~\cite{reid_ultrafiltration_1960} The most limiting 
  drawback to CA membranes is their chemical instability, undergoing severe 
  hydrolysis at pH values outside the range 4--6.~\cite{vos_kinetic_1966}

  Amorphous thin film composite (TFC) polyamide membranes are the current industry 
  standard for high purity RO separations.~\cite{baker_membrane_2012} Developed 
  in the 1980s, polyamide membranes are capable of rejecting salt to a higher 
  degree and can operate at wider pH ranges and higher feed pressures than 
  CA membranes.~\cite{cadotte_interfacial_1981} TFC polyamide membranes are 
  typically formed by the interfacial polymerization of a polyamide selective 
  layer on the surface of a microporous polysulfone support layer. The support
  layer gives the membrane mechanical strength without hindering the flow of 
  water thereby permitting a very thin selective layer where molecular 
  separations occur.~\cite{kucera_biofouling_2019}

  The dense polyamide selective layers of TFC membranes require high feed pressures,
  in the range of 5 -- 120 bar, in order to achieve a useful permeate flux.~\cite{van_der_bruggen_review_2003}
  While the energy requirements to attain these pressures are still less than 
  the energy requirements of thermal distillation\cite{morton_environmental_1997}, it 
  is often desirable to create membranes that can operate at lower feed pressures.
    %BJC2: probably could use some citations here
  It is not always necessary to achieve the high purity permeate streams
  resulting from RO. It is also possible that a target separation may be possible with 
  membranes containing larger pores. This has driven research into nanofiltration.

@@

  Since their introduction in the late 1980s, NF membranes have been applied towards
  an increasing number of separations since they bridge the gap in capabilities
  between reverse osmosis and ultrafiltration.~\cite{eriksson_nanofiltration_1988}
  Nanofiltration membranes have pores on the order of 1 nm in size. Larger pore sizes
  allow lower operating pressures (3 -- 20 bar) which lowers energy consumption.~\cite{van_der_bruggen_review_2003}
  An ideal NF membrane should have densely packed, uniform-sized and non-tortuous 
  pores.~\cite{jackson_nanoporous_2010} This combination has been practically difficult to realize.

  The most widely used methods of NF membrane fabrication are interfacial polymerization
  and phase inversion. NF membranes created by interfacial polymerization are made similarly
  to RO membranes. By modifying the monomer materials, one can control the resultant
  membrane's structural morphology in order to create pores larger than those in RO 
  membranes.~\cite{seman_nanofiltration_2010} During phase inversion, one submerges a 
  polymer, dissolved in a solvent, in a non-solvent. A solid, porous polymer membrane 
  is all that remains once all solvent has been removed by non-solvent exchange.~\cite{smolders_microstructures_1992}
  Although commercially dominant, both of these techniques suffer from tortuous pores
  and polydisperse pore sizes which results in far from ideal selectivities.

  Track-etching is another NF fabrication technique which attempts to eliminate 
  polydispersity and tortuosity. During track-etching, a polymer film is bombarded with 
  charged particles, then chemically etched to create pores.~\cite{apel_track_2001}
  Although the pores are uniform and non-tortuous, benefiting selectivity, the 
  membranes have an extremely low pore density and subsequently low permeability.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{figs/graphene.pdf}
  \caption{graphene}\label{fig:graphene}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{figs/cnt.png}
  \caption{carbon nanotube}\label{fig:cnt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} % https://america.iza-structure.org/IZA-SC/framework_3d.php?STC=LTA
  \includegraphics[width=\textwidth]{figs/zeolite.png}
  \caption{zeolite, ZK-4}\label{fig:zeolite}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{figs/mof.png}
  \caption{metal organic framework, MOF-5}\label{fig:mof}
  \end{subfigure}
  \caption{(a) Pristine graphene (left) and graphene oxide (right) are atomically thick,
  offering the potential for high permeabilities. (b) Carbon nanotubes exhibit extemely
  fast water transport due to their hydrophobic interiors. (c) Zeolites, like ZK-4 
  shown here, possess cage-like structures that can trap solutes. (d) Metal organic
  frameworks are flexible, porous structures capable of selective adsorption. }\label{fig:nanostructured_materials}
  \end{figure}

  \textit{Graphene and graphene oxide}: Ultrathin-film graphene and graphene oxide membranes
  are an active area of research because they are atomically thick materials and therefore 
  offer potential for extremely high permeability membranes.~\cite{humplik_nanostructured_2011} 
  The tightly packed aromatic network of $sp^2$ hybridized carbon atoms that make up 
  graphene sheets precludes the passage of even the smallest monoatomic species, helium.~\cite{bunch_impermeable_2008}
  Controllably-sized pores have been created in pristine graphene by plasma etching
  and by charged particle bombardment.~\cite{surwade_water_2015,russo_atom-by-atom_2012}
  Graphene oxide (GO), a highly oxidized modification of graphene, consist of epoxy,
  hydroxyl and carboxylate functional groups.~\cite{hegab_graphene_2015}
  Many groups study GO membranes because they have improved antibacterial properties 
  over graphene, leading to less membrane fouling, and because their functionality can be
  modified in order to achieve solute-specific separations.~\cite{liu_antibacterial_2011,he_bioinspired_2013}

  Multilayered GO membranes have shown considerably more progress than their single-layered
  counterparts because their synthesis procedures are low cost and scalable.~\cite{zhang_size-controlled_2009} 
  Despite their promise, scalable synthesis of monolayer graphene and GO membranes without
  introducing microscopic, performance degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018} GO sheets are particularly easy to make into 
  multilayered membranes since they stack on top of each other,
  held together by hydrogen bonds.~\cite{homaeigohar_graphene_2017} The potential of 
  graphene and GO membrane certainly merits their continued exploration. Many groups are still
  working in order to work on better control and standardization of their synthesis, improving
  their designability for specific separations, and further improving their scalability.~\cite{wei_multilayered_2018}

  \textit{Carbon nanotubes}: Carbon nanotubes (CNTs) have shown promise as aqueous 
  separations membranes due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011}
  CNTs are an allotrope of carbon consisting of nested graphitic cylinders.~\cite{iijima_helical_1991}
  Functionalization of the CNT pore entrances can potentially lead to highly selective 
  separations.~\cite{fornasiero_ion_2008} One can control CNT diameters down to 0.5 nm 
  in size which has inspired the creation of CNT membranes for reverse osmosis 
  applications.~\cite{sun_creating_2000}

  Molecular dynamics simulations have predicted and experiments have confirmed that 
  CNTs are extremely fast transporters of water.~\cite{hummer_water_2001,joseph_why_2008}
  MD simulations by Kalra \textit{et al.} predicted water transport at rates comparable
  to the transmembrane protein aquaporin-1.~\cite{kalra_osmotic_2003} Experiments have 
  exhibited water transport rates orders of magnitude higher than predictions by 
  continuum modeling.~\cite{holt_fast_2006} This enhancement has been explained as a 
  result of nearly friction-less single-file water flow through the hydrophobic CNT 
  interiors.~\cite{kalra_osmotic_2003}

  Practically, dispersing and aligning CNTs into a polymer matrix is extremely
  difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
  Some work has been to functionalize the carbon nanotubes in order to better incorporate 
  them, however facile incorporation at scale remains a challenge yet to be overcome.~\cite{kim_polysulfone_2007}

  \textit{Zeolites}: Zeolite membranes offer the potential for permeabilities 
  comparable to ultrafiltration, with selectivities as good as NF and RO.~\cite{hoek_nanotechnology-based_2014}
  Zeolites have highly uniform nm-sized crystalline structures with cage-like
  cavities that allow movement and trapping of small solutes. The crystalline 
  frameworks are typically formed by networks of silicon and/or aluminum each 
  attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
  One can replace the silicon and aluminum atoms via ion exchange in order
  to control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}

  A number of studies have tested the permeability and sodium salt rejection
  of various zeolite membranes, however their high permeabilities typcially come
  at the cost of low selectivities and vice versa.~\cite{li_desalination_2004,duke_seawater_2009}
  Zeolite-coated ceramic membranes have shown the most promise for RO applications
  because of their mechanical stability under high applied pressure and chemical
  stability for withstanding disinfectants.~\cite{pendergast_review_2011}
  Unfortunately, zeolites are prone to defects in their crystalline structure which 
  would require substantial modification to the traditional synthesis procedures
  in order to fix.~\cite{kumakiri_application_2000}

  \textit{Metal organic frameworks}: Finally, metal organic frameworks (MOFs) have been
  recently introduced to the field of aqueous separations. MOFs are highly porous and 
  flexible crystalline structures made of metal clusters coordinated to organic 
  ligands.~\cite{furukawa_chemistry_2013} Purification of waste streams using MOFs 
  are typically achieved by selective adsorption or catalytic degradation.~\cite{li_metalorganic_2012}
  There is a much wider breadth of research into MOFs as gas separation membranes 
  because they tend to be unstable in water.~\cite{feng_construction_2013} But a new
  generation of water-stable MOFs is making it possible to use them in 
  aqueous separations.~\cite{wang_ultrastable_2015}

  Recently, MOFs have been applied towards the removal of a number of common water 
  contaminants including dyes, pesticides, endocrine disruptors, and pharmaceuticals.
  \cite{zhang_fabrication_2015,seo_adsorptive_2015,li_metalorganic_2015,hasan_adsorptive_2012}
  Although aqueous MOF separations are theorized to surpass the performance 
  of conventional membrane separations, it is still a budding field. The stability and
  reusability of the material must be guaranteed before they can be implemented
  at scale.~\cite{dias_towards_2015}

@@

  Amphiphilic molecules that self-assemble to form ordered nanostructures have shown
  potential as nanoporous membranes. People are particularly interested in 
  self-assembled materials because they offer the potential of uniform-sized pores
  and tunable chemical and physical properties.~\cite{werber_materials_2016}
  There are two major classes of self-assembled amphiphilic materials being studied
  for separation applications: block copolymers and liquid crystals.

  Block copolymers contain more than one distinct repeating unit which facilitates 
  the formation of self-assembled structures based on differences in the subunit 
  amphiphilicity.~\cite{jackson_nanoporous_2010} Block copolymer membranes typically 
  have pores on the order of 5-50 nm which makes them useful for UF applications, 
  however there is progress being made towards smaller pores in the NF regime.~\cite{gu_tailoring_2015}
